|                                                              |
|--------------------------------------------------------------|
| title: "Classification Analysis on Credit Card Applications" |
| output: html_document                                        |
| #date: "`r Sys.Date()`"                                      |

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 2.1

In the health insurance industry we aim to classify members that are at risk for high cost and high healthcare utilization. Risk scores can be used to identify potential cost saving measures, targeted interventions, or predict healthcare expenses among other use cases. These are 5 predictors to help classify members into risk categories.

1.  Demographics factors such as location, age, gender
2.  Socionomic factors such as income level
3.  History of previous healthcare
4.  Lifestyle factors
5.  Insurance policy details

## Question 2.2.1

**Import Packages and IDE Prep**

```{r}
#rm(list = ls())
set.seed(2)
library(kernlab, plyr)
library(kknn)
options(scipen=999)
```

**Read in Dataset(s)**

```{r}
data_with_header =
  read.table(
    '/Users/akashhingu/Documents/Georgia Tech/ISYE 6501/Week 1/HWK/credit_card_data-headers.txt'
    ,header = TRUE
  )
head(data_with_header,5)
```

\newpage

**Create KSVM Model as function that takes in parameter C and returns accuracy**

```{r, tidy=TRUE}
## creating a function for ksvm model that takes C as variable and returns accuracy
ksvm_model_by_c = function(i) {
  pred = list()
  model = ksvm(
    as.matrix(data_with_header[, 1:10]),
    as.factor(data_with_header[, 11]),
    type = "C-svc",
    kernel = "vanilladot",
    C = i,
    scaled = TRUE
  )
  # see what the model predicts
  pred <- predict(model, data_with_header[, 1:10])   
  # see what fraction of the model’s predictions match the actual's
  sum_prediction = sum(pred == data_with_header[, 11]) / nrow(data_with_header) 
  return(sum_prediction)
}

ksvm_model_by_c(100) ##call ksvm model using C=100 ##accuracy

```

**Create for loop to try different values of C**

```{r}
# empty data frame to eventually store C values with their model accuracy predictions
results_df = data.frame(C = numeric(), model_accuracy = numeric()) 
for (i in c(0.1, 1,10,100,1000,10000,100000)) {
  # loop through sequence for different orders of magnitude of C values
  model_accuracy = ksvm_model_by_c(i) # call ksvm_model_by_c function
  #append results to empty dataframe
  results_df = rbind(results_df, data.frame(C = i, model_accuracy = model_accuracy)) 
}
#sort results df by highest sum of predictions
results_df[order(results_df$model_accuracy, decreasing = TRUE), ] 

```

\newpage

**Calculate KSVM Model Formula**

```{r}
  model = ksvm(
    as.matrix(data_with_header[, 1:10]),
    as.factor(data_with_header[, 11]),
    type = "C-svc",
    kernel = "vanilladot",
    C = 100, #using 100 as C
    scaled = TRUE
  )

# calculate a1...am
a = colSums(model@xmatrix[[1]] * model@coef[[1]])
a


# calculate a0
a0 = -model@b
cat("a0 = ",a0)

# calculate accuracy of model
pred <- predict(model, data_with_header[, 1:10])   
sum_prediction = sum(pred == data_with_header[, 11]) / nrow(data_with_header)
cat("Model Accuracy: ",sum_prediction)


```

**Classifier Equation:**

(-0.0010065348 x A1) + (-0.0011729048 x A2) + (-0.0016261967 x A3) + (0.0030064203 x A8) + (1.0049405641 x A9) + (-0.0028259432 x A10) + (0.0002600295 x A11) + (-0.0005349551 x A12)\
+ (-0.0012283758 x A14) + (0.1063633995 x A15) + 0.08158492 = 0

This model has a 86.39% accuracy

\newpage

## Question 2.2.3

**Looping Through KKNN Model**

```{r}
#create empty data frame to store values
kknn_results_df = data.frame(k = numeric(), 
                   prediction_kknn = numeric(), 
                  actual_kknn = numeric()) 
# for loop to test different K values by each data point
for (k in seq(1,25,1)) {
  for (p in seq(1,nrow(data_with_header))){
    kknn_model = kknn(R1~.
                      ,data_with_header[-p,]
                      ,data_with_header[p,]
                      ,k = k
                      ,kernel = "optimal"
                      ,distance = 2
                      ,scale = TRUE
    )
    prediction_kknn = round(kknn_model$fitted.values) #prediction
    actual_kknn = data_with_header[p,11] #actual
    kknn_results_df = rbind(kknn_results_df, data.frame(k = k, prediction_kknn =                      prediction_kknn, actual_kknn = actual_kknn)) #append to data frame
  }

  
}
#Viewing new dataframe which contains K, predictions, and actuals from KKNN model
head(kknn_results_df)
tail(kknn_results_df)

```

```{r}
#create empty data frame to store values
accuracy_kknn_df_by_k = data.frame(k = numeric(), acc_kknn_by_k = numeric())

#loop through K values and calculate and average accuracies for different K values from previous data frame
for (k in unique(kknn_results_df$k)) {
  k_values = kknn_results_df[kknn_results_df$k == k, ]
  acc_kknn_by_k = sum(k_values$prediction_kknn == data_with_header[, 11]) /
  nrow(data_with_header)
  accuracy_kknn_df_by_k = rbind(accuracy_kknn_df_by_k,
  data.frame(k = k, acc_kknn_by_k = acc_kknn_by_k))
  
}

#sort results df by highest sum of predictions
accuracy_kknn_df_by_k[order(accuracy_kknn_df_by_k$acc_kknn_by_k, decreasing = TRUE), ] 

```

**Based on the accuracy averages for the different k's choosing K = 12 will yield the highest model accuracy at 85.32%**


## Question 3.1 a

**Cross validation using Leave One Out Cross Validation (LOOCV)**

```{r}
#3.1 a -- leave one out cross validation using train.kknn()



## LOOCV on train dataset
kknn_LOOCV = train.kknn(
  R1~.,
  data = data_with_header,
  kmax = 100,
  distance = 2,
  kernal = 'optimal',
  scale = TRUE
)

results_df_kknn_loocv = data.frame(k = numeric(), acc = numeric()) #empty dataframe

#loop through k:kmax to calculate model accuracies
for (k in unique(seq(1,100,1))) {
  prediction = round(kknn_LOOCV$fitted.values[[k]])
  acc = sum(prediction == data_with_header[,11])/nrow(data_with_header)
  results_df_kknn_loocv = rbind(results_df_kknn_loocv, data.frame(k = k, acc = acc)) 
  
}


#dataframe ordered by acc desc
results_df_kknn_loocv[order(results_df_kknn_loocv$acc, decreasing = TRUE),][1:10,]

```

**Using the LOOVC we find the kknn model with k = 12 has one the best accuracy at 85.32%**

## Question 3.1 b

```{r}
#3.1 b

##build train, validation, and test data splits 70-15-15 splits

train70 = sample(nrow(data_with_header),size = nrow(data_with_header)*.70)
train_df = data_with_header[train70,]

valid_test_df = data_with_header[-train70,]
valid_test_split = sample(nrow(valid_test_df), size = nrow(valid_test_df)*.5)
valid_df = valid_test_df[valid_test_split,]
test_df = valid_test_df[-valid_test_split,]


#check df sizes
nrow(data_with_header)
nrow(train_df)/nrow(data_with_header)
nrow(valid_df)/nrow(data_with_header)
nrow(test_df)/nrow(data_with_header)

##empty df
results_df_kknn_train = data.frame(k = numeric(), acc = numeric()) 
#Loop through different K's
for (k in seq(1,100,1)) {
  kknn_model = kknn(R1~.
                    ,train_df
                    ,valid_df
                    ,k = k
                    ,kernel = "optimal"
                    ,distance = 2
                    ,scale = TRUE
  )
  prediction = round(kknn_model$fitted.values)
  acc = sum(prediction == valid_df[,11])/nrow(valid_df)
  results_df_kknn_train = rbind(results_df_kknn_train, data.frame(k = k, acc = acc)) 
  
}

results_df_kknn_train[order(results_df_kknn_train$acc, decreasing = TRUE),][1:10,]


```

**K = 23 has highest accuracy at 87.8%**

**Use trained and validated model on test df at k = 23**

```{r}
kknn_model_tested = kknn(R1~.
                  ,train_df
                  ,test_df
                  ,k = 23
                  ,kernel = "optimal"
                  ,distance = 2
                  ,scale = TRUE
)

prediction = round(kknn_model_tested$fitted.values)
acc = sum(prediction == test_df[,11])/nrow(test_df)
acc


```

**KKNN Model (with k = 23) produces best model with accuracy at 82% on test data**
